|

### 2021-08-02 dissertation/msc_dissertation.Rmd
Minor viz adjustments and further thesis writing

# Introduction
     The Covid-19 pandemic has reaffirmed the need for rigours scientific research to inform decision-making on a scientific and societal level. However, particularly the social sciences have a history of inflated positive findings (Ionanidis) leading to the reproducibility crisis [@pashler2012]. @simmons2011 demonstrated how the so-called *researcher degree of freedom* affect study results. This concept refers to the analytical choices researchers make and their induced variation in results. For instance, sample size can be highly deterministic of finding statistically significant effects. @john2012 found 70% of researchers have at least once stopped their data collection based on a preliminary analysis. P-value simulations as a function of sample size showed an increased likelihood of obtaining false-positives for small sample sizes. Specifically, a sample size of ten had a 22.1% chance of being a false positive. Adding 20 observations per condition decreased the probability to 12.7% [@simmons2011]. The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish. To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers, respectively.

 @silberzahn2018 investigated the effects of analytical choices on results. Specifically, they had 29 teams of researchers investigate the same research question with the same dataset. Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g. wanting/needing to publish). The variation between the analytic approaches was substantial. There were 29 different analyses with 21 different combinations of covariates. Twenty teams found a statistically significant effect, and the odds-ratios ranged from 0.89 to 2.93 (median = 1.31). The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results.

This study puts (psychological) research in a delicate position, because if researchers with honest intentions come to such drastically different conclusions despite starting off on the identical dataset what does this mean for the rest of studies?

@patel2015

# Methods

# Results

# Discussion

>>>>

# Introduction
     The Covid-19 pandemic has reaffirmed the need for rigours scientific research to inform decision-making on a scientific and societal level. However, particularly the social sciences have a history of inflated positive findings (Ionanidis) leading to the reproducibility crisis [@pashler2012]. @simmons2011 demonstrated how the so-called *researcher degree of freedom* (RDF) affect study results. This concept refers to the analytical choices researchers make and their induced variation in results. For instance, sample size can be highly deterministic of finding statistically significant effects. @john2012 found 70% of researchers have at least once stopped their data collection based on a preliminary analysis. P-value simulations as a function of sample size showed an increased likelihood of obtaining false-positives for small sample sizes. Specifically, a sample size of ten had a 22.1% chance of being a false positive. Adding 20 observations per condition decreased the probability to 12.7% [@simmons2011]. The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish. This example shows how easy it is to engineer statistically significant effects. To prevent such bad practices @simmons2011 suggested requirements and guidelines for authors and reviewers.

 Among these requirements and guidelines are a minimum sample size, reporting all collected variables and presenting results with and without covariates [@simmons2011]. These are important suggestions, but they struggle to cope with the full spectrum of the analytical possibilities. @silberzahn2018 further scrutinised the effects of RDF. Specifically, they had 29 teams of researchers investigate the same research question based on the same dataset. Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g. wanting/needing to publish). The variation between the analytic approaches was substantial. There were 29 different analyses with 21 different combinations of covariates. Twenty teams found a statistically significant effect; the odds ratios ranged from 0.89 to 2.93 (median = 1.31). The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results. Similar studies support these RDF effects [@botvinik-nezer2020]. These studies show that even extended efforts do not necessarily yield conclusive results. Moreover, this crowdsourcing approach is not practical.

 @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis, or specification-curve analysis, as an alternative to crowdsourcing approaches. This approach requires a researcher, ideally a more than one, to conduct with every plausible analysis, aggregate and report it, and conduct an inferential test across all possibilities to come to a conclusion. "Plausible" refers to statistically valid tests appropriate to the research question and non-redundancy [@simonsohn2020]. This approach is statistically more complex and computationally more intense than conventional analyses, but makes inherent strategy bias and noise transparent, and forces researchers to look at the entire "garden of forking paths" [@gelman2013]. A good example of quantifying the extend of variation due to analytical choices was conducted by @patel2015. They tested the association of clinical, environmental and physiological variables with all-cause mortality. Specifically, they specified three model approaches (gaussian, cox and binomial) and 13 different covariates. Their algorithm ran every possible covariate combination (on top of the base variables) for each model which allowed the authors to assess the "vibration of effect" (VoE). They concluded that a larger VoE calls for more caution when interpreting observational associations.

 Taken together, statistically effects are easy to engineer through adjusting analysis parameters like sample size. Crowdsourcing analysis is a good strategy to understand the RDF induced variation but is impractical due to its substantial personnel, time and effort requirement. Alternatives like multiverse analysis present a promising, low-personnel, but computationally intense alternative to identify RDF induced variation. The present study will build upon @silberzahn2018, @botvinik-nezer2020 and @patel2015. The objective is to conduct a multiverse analysis based on the data used in @silberzahn2018 and identify the VoE i.e., the results space of the outcome variable. This will allow for a better understanding of how robust crowdsourced identified effects are. Due to time constraints this study will function as a prove-of-concept. Specifically, it will take the analytical approach which produced the median outcome and run it with every possible covariate combination. Similarly to @patel2015 this produces a distribution of the outcome effects and its variation provides insight into the @silberzahn2018 median effect robustness. The smaller the variation, the greater is the effect's robustness.
